The evolution of gene expression in the term placenta of viviparous mammals

### 2016-03-25 manuscript/manuscript.Rnw
completely rewrite abstract

\begin{abstractbox}
\begin{abstract}
% \section{Abstract}
  \parttitle{Background} The placenta is arguably the most
  anatomically variable organ in mammals. Understanding gene
  expression patterns among placentas from multiple species can
  illuminate the molecular basis of placenta function as well as
  potentially enrich our knowledge of the biology of obstetrical
  syndromes.

  \parttitle{Results} We measured the expression profiles of the term
  placenta of \Sexpr{star.logs[,length(unique(species))]} species of
  therian (i.e. marsupials and “placental”) mammals using RNA-Seq. We
  identified multiple genes and transcripts which are differentially
  expressed across lineages including prolactin, relaxin, and
  galectins, as well as identifying the non-housekeeping genes which
  are core to placenta function (such as components of annexin
  complexes) in all \Sexpr{star.logs[,length(unique(species))]}
  species studied.

  \parttitle{Conclusions} While there are multiple non-housekeeping
  genes which are highly expressed in all placentas examined and are
  likely to be involved in the basic functions of the mammalian
  placenta, we see incredible lineage-specific diversity in the
  expression and expansion of many genes. These lineage-specific genes
  are likely to play a role in the anatomical variability of the
  placenta and may also be implicated in important human placental
  disorders such as pre-eclampsia.

  \DLA{Complete rewrite abstract to make the points that we have found
    stronger and less weak.}

% We identified significant shifts in gene expression along specific
% evolutionary lineages using a model of character evolution. Genes that
% showed significant elevation in expression along the human lineage
% ($\mathrm{FDR}\le 0.01$) were enriched in GO categories such as
% angiogenesis, tube morphogenesis, cell migration, and regulation of
% inflammatory response. Three of these genes (KISS1, ADAM12, GPC3)
% showed exceptionally high expression in humans; these genes are
% implicated in pre-eclampsia and may indicate that the risk of
% pre-eclampsia recently emerged on the human lineage.

>>>>

\begin{abstractbox}
\begin{abstract}
% \section{Abstract}
  \parttitle{Background} The placenta is arguably the most
  anatomically variable organ in mammals. Understanding gene
  expression patterns among placentas from multiple species
  illuminates the molecular basis of placenta function as well
  obstetrical syndromes affecting humans.

  \parttitle{Results} We measured the expression profiles of
  \Sexpr{star.logs[,length(unique(file))]} term placenta of
  \Sexpr{star.logs[,length(unique(species))]} species of mammals. We
  found lineage-specific differential expression of prolactin,
  relaxin, and galectins as well as increased expression of the
  pre-eclampsia associated genes \textit{ADAM12}, \textit{CRH}, and
  \textit{PAPPA} in humans. We also identified genes which are
  expressed in all placentas, including genes with immune-modulating
  properties (\textit{ANXA2}, \textit{ANXA1}, \textit{S100A11}, and
  \textit{S100A10}) and cell-cell interactions (\textit{LAMC1},
  \textit{LUM}, and \textit{LGALS1}).

  \parttitle{Conclusions} A diverse array of genes are differentially
  expressed and are likely to play a role in the anatomical
  variability of the placenta in the placentas of mammalian lineages.
  However, the expression of a subset of
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes is essential for the proper functioning of the placenta in all
  species analyzed.
